Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-11 ba_resubmit/response_to_reviewers.tex

	\reply{%
\textbf{We thank the Editor, the AE, and the reviewer for their feedback, and pointing out our strengths of the paper and areas that we can improve the quality of our paper. In summary, we have revised our paper in the abstract and introduction as to make our contributions more clear to that of prior work.  Furthermore, we have carefully edited the paper to address issues that were pointed out by the Editorial Board.} \todo
	}

	\newpage

	\pointRaised{E}{%
	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.  Independently, I found many similar problems.  The intersection between what I found and what the AE found is non-empty and so is the symmetric difference.  Also, the AE found some issues that I did not find and vice versa.  So, I am almost certain that an additional review would uncover more problems.  Now, the results presented in the paper seem plausible as do the broad strokes of the derivations, but I must admit that I was not able to follow all the details. This, in large part, was because of the issues that I just mentioned, and I believe that most readers would find it difficult to follow the developments, as these issues are exceedingly distracting.
	}

	\reply{
\textbf{We appreciate these comments, and have done our utmost to address the ones listed by yourself and the AE. Furthermore, we have carefully gone through the paper ourselves, where we have corrected other issues with notation, typos, and various inconsistencies.} \todo
	}

@@

	\pointRaised{E}{%
	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently. For example, $I_{obs()}$ is never defined, $n_{12(Z)}$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.
	}

	\reply{
\textbf{The definition of $I_{obs}()$ now defined. We have defined $n_{12}(Z)$ in the correct place (and removed the error with $D$). We have checked for other notational inconsistencies. Additionally, we have defined $n_{12}(Z)$ when it first appears.}
	}

@@

	\pointRaised{R1}{%
	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where linkages are made within strata of agreement between records across the two databases, and relies heavily on advances made by Sadinle (2017). Overall, the paper presents a practical implementation of a modified Beta Linkage approach to	matching. It demonstrates that even though the new method suffers from the limitations of not seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of features on which to match. Though the novelty of the presented methodology is primarily in the computational approach, I believe it to be a real advance for practical implementation of record linkage in modern applications.

	I have no major concerns about the content or accuracy of the presentation. However, substantial improvements are warranted to confirm my impression. In addition to the recommendation made by Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the explanation of the method should be written more clearly, the simulation comparisons should be repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally, more practical advice would be welcome. All of these points are more specifically described in the itemized list below
	}

	\reply{%
	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017). Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations. We expand on these points further below.} \todo
	}

	\pointRaised{R1}{%
		Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper more easily.}
\reply{%
	We have provided a summary of notation in Appendix 7.1. \textcolor{red}{I'm not sure where I should reference this in the text. Do I need to?}
}

@@

\pointRaised{R1}{%
Section 3, full model: The Phi notation needs to be defined.
}
\reply{%
	\textbf{This has been defined.}
}

\pointRaised{R1}{%
	Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.
}

\reply{%
	\textbf{The expression is now corrected.}
}

\pointRaised{R1}{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…). This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in intermediate conclusions. In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(\gamma_{ij}^f=l)$. This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}

@@

%\reply{%
%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler. I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
%
%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented. \textcolor{red}{I want you guys to look at it though. Its pretty ugly.}
%}

\pointRaised{R1}{%
Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional presented on the first line of page 8.
}

\reply{%
	\textbf{The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior. The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$. We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$. We have updated our explanation to make this more clear.}
}

\pointRaised{R1}{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-$\pi$).
}
\reply{%
	\textbf{This has been corrected.}
}

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
edits

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}.$$
For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics. Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.

@@

	\begin{table}[t]
		\centering
		\begin{tabular}{rrr}
			\hline
			& m & u \\ 
			\hline
			fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
			lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
			day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
			month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
			year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
			\hline
		\end{tabular}
		\caption{Distributions used for $m$ and $u$ probabilties in simulation studies}\label{Tab:distributions}
	\end{table}

@@

>>>>

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$ \textcolor{red}{is a list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$}. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
which is the number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$, and 
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p},$$
which is the number of comparison vectors that exhibit pattern $p$.
For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics. Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.

@@

	\begin{table}[t]
		\centering
		\begin{tabular}{rrr}
			\hline
			& m & u \\ 
			\hline
			fname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
			lname & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{100}, \frac{99}{100}\right)$ \\ 
			day & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{30}, \frac{29}{30}\right)$ \\ 
			month & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
			year & $\left(\frac{19}{20}, \frac{1}{20}\right)$ & $\left(\frac{1}{12}, \frac{11}{12}\right)$ \\ 
			\hline
		\end{tabular}
		\caption{Distributions used for $m$ and $u$ probabilities in simulation studies}\label{Tab:distributions}
	\end{table}

@@

	\hypertarget{app:glossary}{%
	\subsection{Summary of Notation}\label{app:glossary}}

		\begin{table}[h!]
		\centering
		\begin{tabular}[h!]{ll}
			\hline
			Symbol & Description \\
			\hline
			$\bm{X}_1, \bm{X}_2$ & datasets\\
			$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
			$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
			$f \in 1, \ldots F$ & index over fields used for comparisons \\
			$l \in 1, \ldots L_f$ & index over agreement levels for field $f$ \\
			$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
			$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
			$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
			$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
			$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
			$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
			$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
			$r_{j_p}$ & list of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
			$h_p$ & one hot encoding of agreement pattern $p$ \\
			$H_{j_p}$ & number of records in $\bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ \\
			$H_p$ & number of comparison vectors that exhibit pattern $p$ \\
			\hline
		\end{tabular}
	\end{table}
